---
title: "Trophic Position and Pollutant Concentration in Ringed Seals"
Authors: "Kaiulani Sund and Sarah Godwin"
output: html_notebook
---

## Description 
Our study took place in four regions in the Canadian Arctic. Scientists measured the amount Persistent Organic Pollutants (POPs) in ringed seals. We are comparing Trophic Position and pollution concentration to see if there is a correlation, under the hypothesis that individuals foraging at a higher trophic level will have higher pollutant concentrations as a result of biomagnification. We focused on the pollutants PCB and DDT. 

The dataset we used can be found here: https://open.canada.ca/data/en/dataset/d0613682-0862-45d8-a568-88d947f086e5



## Installing and Loading Packages
```{r}
install.packages("tidyverse") #installing tidyverse package
library("tidyverse") #loads the package
install.packages("rstatix") #installing rstatix package
library("rstatix") #loads the package
```


## Data Loading
Seals data is put into a tibble table and we change N/A blocks to numeric values and change character columns to numeric values.
```{r}
Seals <- tibble(read.csv(file.choose(), header = TRUE)) #uploads the data as a tibble, header = True keeps the column title

Seals = 
  Seals %>% #choosing data set
  mutate(TrophicPosition = as.numeric(as.character(TrophicPosition))) %>% #Changes N/A to a numeric value NA
  mutate(Sum_DDTs = as.numeric(ifelse(Sum_DDTs == "N/A", "NA", as.numeric(as.character(Sum_DDTs))))) %>% #Changes N/A to a numeric value NA
  mutate(Sum_PCBs = as.numeric(ifelse(Sum_PCBs == "N/A", "NA", as.numeric(as.character(Sum_PCBs))))) #Changes N/A to a numeric value NA 

#you will get an error but the code works - it is saying there were character values (N/A) that were made into true NAs

```


## Descriptive Statistics
In this section, we calculate the mean, max, and minimum value for each of the parameters of interest. 
```{r}
summary(Seals) #summarizes all columns in Seals Data - note that several additional columns would need to be made numeric before calculating meaningful summary statistics 

Seals %>% #choosing data set
summarize(meanTrophic = mean(TrophicPosition, na.rm = TRUE), maxTrophic = max(TrophicPosition, na.rm = TRUE), minTrophic = min(TrophicPosition, na.rm = TRUE)) #taking mean, max and min values of TrophicPostition na.rm removes N/A values

Seals %>% #choosing data set
summarize(meanPCB = mean(Sum_PCBs, na.rm = TRUE), maxPCB = max(Sum_PCBs, na.rm = TRUE), minPCB = min(Sum_PCBs, na.rm = TRUE)) #taking mean, max and min values of PCBs na.rm removes N/A values

Seals %>% #choosing data set
summarize(meanDDT = mean(Sum_DDTs, na.rm = TRUE), maxDDT = max(Sum_DDTs, na.rm = TRUE), minDDT = min(Sum_DDTs, na.rm = TRUE)) #taking mean, max and min values of DDTs na.rm removes N/A values
```


## Histogram
Creating a Histogram to visualize the number of seals at each trophic level.
```{r}
Seals %>% #choosing data set
ggplot(aes(x=TrophicPosition)) + geom_histogram(bins = 30) + #making a histogram showing the number of seals at each Trophic Position
theme_bw() + #removing gray background
ylab("number of individuals") #changing the name of the y axis
```


## Linear Regression Models
**PCBs and Trophic Position**
Creating a linear regression model to determine if there is a correlation between Trophic Position and PCB levels.
```{r}
Seals %>%
ggplot(aes(x= TrophicPosition, y = Sum_PCBs)) + geom_point() + geom_smooth(method = "lm") + #makes linear regression, with data points and line of best fit
theme_bw() + #removes gray background
xlab("Trophic Position") + #changing the name of the x axis
ylab("Total PCB Level (ng/g l.w.)") + #changing the name of the y axis #units are nanograms per gram lipid weight
ylim(0, 4000) + #limits y axis to a range of 0 to 4000 #Note - although there are some PCB values above 4000, they do not have trophic position information  
  theme(axis.line = element_line(colour = "black"),
  panel.grid.major = element_blank(), #removes major gridlines
  panel.grid.minor = element_blank()) + #removes minor gridlines
  theme(axis.text.x = element_text(color = "black", size = 14)) + #changes color and size of x axis labels
  theme(axis.title.x = element_text(face = "bold", color = "black", size = 14)) + #changes color and size of x axis title
  theme(axis.text.y = element_text(color = "black", size = 14)) + #changes color and size of y axis labels 
  theme(axis.title.y = element_text(face = "bold", color = "black", size = 14))  #changes color and size of y axis title


summary(lm(Sum_PCBs ~ TrophicPosition, data = Seals)) #Are PCB Level and Trophic Position correlated

#Result: There is a significant positive correlation between PCB levels and trophic position (p < 0.001)
```

**DDTs and Trophic Position**
Creating a linear regression model to determine if there is a correlation between Trophic Position and DDT levels.
```{r}
lm(Sum_DDTs ~ TrophicPosition, data = Seals) %>% ##taking DDT and Trophic Position from Seals data for graph
ggplot(aes(x= TrophicPosition, y = Sum_DDTs)) + geom_point() + geom_smooth(method = "lm") + #makes linear regression, with data points and line of best fit
theme_bw() + #removes gray background
xlab("Trophic Position") + #changing the name of the x axis
ylab("Total DDT Level (ng/g l.w.)") + #changing the name of the y axis ##units are nanograms per gram lipid weight
  theme(axis.line = element_line(colour = "black"), 
  panel.grid.major = element_blank(), #removes major gridlines
  panel.grid.minor = element_blank()) + #removes minor gridlines
  theme(axis.text.x = element_text(color = "black", size = 14)) + #changes color and size of x axis labels
  theme(axis.title.x = element_text(face = "bold", color = "black", size = 14)) + #changes color and size of x axis title
  theme(axis.text.y = element_text(color = "black", size = 14)) + #changes color and size of y axis labels 
  theme(axis.title.y = element_text(face = "bold", color = "black", size = 14))  #changes color and size of y axis title



summary(lm(Sum_DDTs ~ TrophicPosition, data = Seals)) #Are DDT Level and Trophic Position correlated
#Result: There is a significant positive correlation between DDT levels and trophic position (p < 0.001)
```


**DDTs and PCBs**
Creating a linear regression model to determine if there is a correlation between PCB levels and DDT levels.
```{r}
Seals %>%
ggplot(aes(x= Sum_PCBs, y = Sum_DDTs)) + geom_point() + geom_smooth(method = "lm") + #makes linear regression, with data points and line of best fit
theme_bw() + #removes gray background
xlab("Total PCB Level") + #changing the name of the x axis
ylab("Total DDT Level")  + #changing the name of the y axis
ylim(0, 6000) + #limits x axis to a range of 0 to 6000
xlim(0, 6000) #limits y axis to a range of 0 to 6000

summary(lm(Sum_DDTs ~ Sum_PCBs, data = Seals)) #Are DDT Level and Trophic Position correlated

#Result: There is a significant positive correlation between PCB levels and DDT levels (p < 0.001)
```
